---
title: "rView"
output: github_document
---

This is a package with a single function that uses the server used by the 'rmote' package
to send data objects over the network to be viewed using the RStudio viewer in a local session.

## Setting up / Installing packages

In order to use this package you must first install two packages, `servr` and `rmote`, into your _remote_ R library (i.e., the one on the compute cluster).  `servr` is on CRAN and can be installed with:
be installed with:
```{r, eval=FALSE}
install.packages("servr")
```
`rmote` is not on CRAN (and may not ever be), so you must install it with:
```{r, eval=FALSE}
remotes::install_github("cloudyr/rmote")
```
There might be a few dependencies that are not automatically installed, and so it might take several
round of installation to finally get those two packages installed.  Once they are installed, however,
you can then install this 'rView' package on both your local and remote computers, by doing this on both of them:
```{r, eval=FALSE}
remotes::install_github("eriqande/rView")
```


## Initiating rmote and rView on a remote computer cluster

Here is what a session looks like on a compute cluster. Some of this
is a workaround because the compute nodes may not be free to do port forwarding, but the
head node(s) typically can.

1. _From RStudio terminal_, login to the head node on the cluster
    a. If you would like, establish or attach to a tmux or screen session. If you don't know that that
    means, then 
2. _From RStudio terminal_ get a shell on compute node (`srun --pty /bin/bash` of `sinteractive`, etc.,
depending on your cluster), and _start an R session by typing `R` at the command line of the compute node.
3. In the R session on the remote machine, start the 'rmote' server, but make sure that it is writing to a directory
in your home directory. (This is necessary so we can relay the contents of that directory from
a port on the head node. The defaut 'rmote' directory location in `/tmp` is accessible only
by the compute node, not the head node.) This directory can be on any storage that is shared by the different
nodes in the cluster. (So, for example, you could put it in your `scratch` area, if desired.  For example,
for me, I might use, the directory `/home/eanderson/rmote_server`.  Also you need to use a port number.  The
default port number is 4321, but you should choose a different one to make sure you don't collide with another
user.  Here I will use 4111, but you should use a different number, let us say in the range 4000 to 5000.
    ```{sh, eval=FALSE}
    library(rmote)
    start_rmote(server_dir = "/home/eanderson/rmote_server", port = 4111)
    ```
4. Once you have started the 'rmote' web server from your remote R session, you can then load the 'rView' package with:
    ```{r, eval=FALSE}
    library(rView)
    ```
5. Now, back on your local machine, in a separate terminal (it does not need to be in
RStudio...you could use the Terminal app, or
iTerm on a Mac), login to the head node, while enabling port forwarding from your port number (4111 in the
example) on the head node to the same port number on your local computer.  That looks like this
    ```{sh, eval=FALSE}
    ssh -L 4111:localhost:4111 eanderson@sedna.nwfsc2.noaa.gov
    
    # and once you get that connection, you must start R on the head node and give the command
    rmote::start_rmote(server_dir = "/home/eanderson/rmote_server", port = 4111)
    ```
This starts R on the head node, and within that R session a small web server is run that
serves up data found in the `server_dir` on port 4111.  This is not a compute-intensive job,
so the sysadmins should not be upset that this is running on the head node.
6. Also, back on your local machine, open a browser window (like Chrome or Safari) and
go to the address of the port number on your local machine. In our example, the
port is 4111 so you must go to: `http://localhost:4111/`.  This
is where `rmote` will serve up plots and help screens.


## Using rmote

This is very straightforward.  The 'rmote' package, when invoked on the remote machine,
overwrites the standard `print` function for base plots, ggplots, and the help function.
So, if you print a ggplot2 object, rather than trying to print on the remote machine and finding
nowhere to print it, when you are running 'rmote', the object will be printed to the
`server_dir` location and served to your local web browser via ssh port forwarding.  You can try it like this:

First, do:
```r
?rmote::start_rmote
```
on the remote machine; then check you web browser that is pointed to  `http://localhost:4111/`.

Second, you might try printing a ggplot object in the R session on the remote machine:
```r
library(tidyverse)
tibble(x = 1:10, y = (1:10)^2) %>%
  ggplot(aes(x = x, y = y)) + 
  geom_point(colour = "blue", size = 2) + 
  geom_line()
```
That will cause the plot to be printed to the web page on the browser. 

## Using rView

When all is set up as above, we can send objects to our local RStudio data viewer like this:

* First way, by object name:
```r
# on remote R session
rView(mtcars)

# then, on local session
rView(mtcars)

```

* If the object to be view is piped in, then do like this:
```r
# on remote R session
library(tidyverse)
diamonds %>%
  filter(cut == "Ideal") %>%
  rView()
  
# then, on local session get it by:
rView(.)

# or by piping anything into rView()...
```

### The `rview_pass_through` option

There might be times when
you would like to add `rView()` commands to the actual R code blocks in an RMarkdown document.
For example, if you are working on an RMarkdown document, then you might have an R code black
that creates a new data object (a tibble maybe) called `my_result`.  To print that to the RMarkdown document, you
might end the code block with a line that says:
```r
my_result
```
Doing so would cause the `my_result` tibble to be printed in the RMarkdown document.

If, instead, you wrote in that code block:
```r
rView(my_result)
```
then, it would be quite easy, when working through the code blocks on a remote machine
(and the local machine) to send that code to the remote and the local R interpreters
in order to view the remote data object on the local RStudio data object browser.  However,
what if you actually just want to knit the RMarkdown document?  That is where the
`rview_pass_through` option comes into play.

If you set that option to TRUE, with:
```r
options(rview_pass_through = TRUE)
```
Then the `rView` commands will be ignored, and effectively it will just default-print the
object in the console/RMarkdown document.

You can restore the regular behavior of `rView` with:
```r
options(rview_pass_through = NULL)
```


## Making the initialization process quicker

To keep from having to type the `start_rmote` command multiple times, you can put that command
(with your preferred port) in your `.Rprofile`.  Then you would always be in `rmote` mode
when you start R on your remote machine.  But then that could screw you up if you are running
R in batch mode.  So it is probably better to just define a function in your `.Rpofile`
that is easy to type, like:
```r
rmote <- function(server_dir = "~/rmote_server", port = 4949) {
  rmote::start_rmote(server_dir = server_dir, port = port)
}
```
that includes the server_dir and the port that you will always want to use.  Then, 
when you enter R, you just have to type `rmote()`, but if you wanted to
change the server_dir and the port, you still could.

